\section{Polynomial Class}
\label{interface-geometry-polynomial}


\noindent
Arbitrary polynomial of order up to $n$ with $d$ parameters can be represented in its expanded form as
\[ p(\vec{u}) = \sum_i A_i \prod_{j = 0}^d u_j^{\mathrm{pow}_{i,j}},  \]
for example in 3D this can be written as
\[ p(\vec{u}) = \sum_i A_i u^{pow_{u,i}} v^{pow_{v,i}} w^{pow_{w, i}},  \]

\noindent
Therefore, we define a Monomial class, which stores a constant multiplier $A$ and powers vector $pow$. \\

\begin{mybox}
\begin{lstlisting}
  PolynomialTraits::Monomial(double prefNew, std::vector<int> powerNew)
  PolynomialTraits::Monomial(double prefNew, int x)
  PolynomialTraits::Monomial(double prefNew, int x, int y)
  PolynomialTraits::Monomial(double prefNew, int x, int y, int z)
\end{lstlisting}
\end{mybox}




\subsection{Methods}
\label{interface-geometry-polynomial-methods}


A polynomial can be constructed either empty, from a single monomial or from a vector of monomials. \\

\begin{mybox}
\begin{lstlisting}
  Polynomial()
  Polynomial(Monomial M)
  Polynomial(std::vector<Monomial> Mvec)
\end{lstlisting}
\end{mybox}

\noindent
The below interface provides methods to perform basic algebraic operations with polynomials and scalars. The method $axpy$ is the scaled addition, equivalent to $this += other * a$ \\

\begin{mybox}
\begin{lstlisting}
  LocalPolynomial & operator+=(const Monomial & otherM)
  LocalPolynomial & operator+=(const LocalPolynomial & other)
  LocalPolynomial & operator*=(const double c)  
  LocalPolynomial & operator*=(const LocalPolynomial & other)  
  void axpy(LocalPolynomial other, double c)
  LocalPolynomial operator+(const LocalPolynomial & other)
  LocalPolynomial operator+(const ctype a)  
  LocalPolynomial operator-(const LocalPolynomial & other)  
  LocalPolynomial operator-(const ctype a)  
  LocalPolynomial operator*(const ctype a)  
  LocalPolynomial operator*(const LocalPolynomial & other)
\end{lstlisting}
\end{mybox}

\noindent
It is possible to differentiate, integrate and evaluate a polynomial. $derivative$ returns the partial derivative of the polynomial wrt coordinate indexed by the parameter. $evaluate$ evaluates the polynomial at the provided local coordinate. $integrateRefSimplex$ integrates the polynomial over the reference simplex of the same dimension as the polynomial, thus returning a scalar. An integral of a monomial over reference simplex has analytical expression, see the proof in appendix \ref{appendix-proof-simplexintegral}  \\


\begin{mybox}
\begin{lstlisting}
  LocalPolynomial derivative(int iDim)
  double evaluate(const LocalCoordinate & point)
  double integrateRefSimplex()
\end{lstlisting}
\end{mybox}


\noindent
The following auxiliary methods can be used to extract additional information about a polynomial. $order$ returns the largest power among all monomials, that is, the sum of powers of that monomial. Magnitude returns the largest absolute value prefactor over all monomials. $to\_string$ converts the polynomial to a string for further text output \\

\begin{mybox}
\begin{lstlisting}
  unsigned int order()
  double magnitude()
  std::string to_string()
\end{lstlisting}
\end{mybox}


\noindent
\uline{Compactify}: adds up all summands with the same power. Sorts the summands by $(x_1,y_1,z_1) < (x_2, y_2, z_2)$, where $x$ has the highest priority and $z$ has the lowest priority. Then all of the repeating powers will be consecutive. Simply loop over sorted polynomial, and to a new polynomial add the sums of all consecutive repeating polynomials.



\subsection{Tests}


\noindent
Currently the tests are only for 1, 2 and 3 dimensions. Most of the tests use intrinsic functionality like polynomial operators and derivatives to construct polynomials and print them to the screen, and request the user to to verify manually if they match the expected polynomials which are also printed. For each dimension there is one test which integrates a non-linear polynomial over simplex and prints out the result which is also compared manually. \\

\textbf{TODO:} These tests can and should be automatized in the future using integer string comparison. The test program should throw an error if a test fails